### 2024-05-27 paper/header.tex
Update paper/header.tex
Co-authored-by: Nadia <76887318+nadiaenh@users.noreply.github.com>

% **************GENERATED FILE, DO NOT EDIT**************

\title{Survey.jl - An Efficient Framework for Analysing Complex Surveys}

\author[1]{Ayush Patnaik}
\affil[1]{XKDR Forum}

\author[2]{Nadia Enhaili}
\affil[2]{Simon Fraser University}

\keywords{Julia, Survey, Statistics, Sampling}

\hypersetup{
pdftitle = {Survey.jl - An Efficient Framework for Analysing Complex Surveys},
pdfsubject = {JuliaCon 2023 Proceedings},
pdfauthor = {Ayush Patnaik},
pdfkeywords = {Julia, Survey, Statistics, Sampling},
}